\LARGE \bf Energy-Aware Dynamic Planning Algorithm for Autonomous UAVs

### 2021-01-20 icra-2021.tex
fixes to the algorithm section

Let us assume that the set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathcal{C}_i\},
\end{equation}
delimits the area where the $i$-th trajectory $\varphi_i$ is free to evolve using the trajectory parameters $c_{i,1},...,c_{i,\rho}$ (the gray area in Figure~\ref{fig:tee1}). $\varphi_i$ is a function of the two coordinates and is equal to zero when a point $\mathbf{p}_k$ is on the trajectory. Physically, this means the UAV is flying exactly over the nominal trajectory. The trajectory parameters allows to change the trajectory. They are a way to alter the nominal trajectory in the initial plan and thus alter the energy signal.
In fact, the algorithm uses the set from Equation~(\ref{eq:area}) to find the optimal trajectory parameters $c_i^0$, such that $\varphi_i$ has the highest instantaneous energy consumption, while still respecting the energy budget. In Figure~\ref{fig:tee1}, the parameter radius of the circle $c_{1,1}$ is replanned (reduced to $c_{1,1}^-$), as e.g., averse atmospheric conditions do not allow to terminate the plan. The sequence of these parameters (control sequence) forms the path $\mathcal{P}$. The gain factors for the trajectory parameters from Equation~(\ref{eq:state-control}) are obtained
\begin{equation}\label{eq:traj-gains}
  \begin{bmatrix}\nu_{i,1} \\ \vdots \\ \nu_{i,\rho}\end{bmatrix}=(\overline{c}_i-\underline{c}_i)^{-1}\rho^{-1}y_{\underline{c}_i}\begin{bmatrix}c_{i,1} \\ \vdots \\ c_{i,\rho}\end{bmatrix},
\end{equation}
where $y_{\underline{c}_i}\in\mathbb{R}$ is the maximum energy consumption of the highest possible parameters level at $i$-th stage. The value is obtained empirically (e.g., using an energy sensor). Whenever the trajectory parameters are not equally distributed, one can define $y_{\underline{c}_i}$ as a diagonal matrix where the items of the diagonal are the highest possible levels of specific trajectory parameters.

We derive the new position $\mathbf{p}_{k+1}$ computing the vector field $\nabla\varphi_i:=\begin{bmatrix}\partial\varphi_i/\partial x & \partial\varphi_i/\partial y\end{bmatrix}^T$, and the direction to follow in the form of velocity vector~\cite{de2017guidance}
\begin{equation*}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varphi_i-k_e\varphi_i\nabla\varphi_i,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation*}
where $E$ specifies the rotation (it influence the tracking direction), and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading $\dot{\mathbf{p}}$ due to the atmospheric interferences, such as wind ($w\in\mathbb{R}$ in the top of Figure~\ref{fig:tee1}).

@@

Let us define $g:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ as the instantaneous computational energy consumption value obtained using the tool. Moreover, let $g(\{\emptyset\})$ be zero.

The gain factors that quantifies the contribution of the computations parameters
\begin{equation}\label{eq:energy-comp}
  \begin{bmatrix}\nu_{i,\rho+1} \\ \vdots \\ \nu_{i,\rho+\sigma}\end{bmatrix}=\begin{bmatrix}c_{i,\rho+1}^{-1} & & \\ & \ddots & \\ & & c_{i,\rho+\sigma}^{-1} \end{bmatrix}\begin{bmatrix}g(c_{i,\rho+1}) \\ \vdots \\ g(c_{i,\rho+\sigma})\end{bmatrix},
\end{equation}
where the matrix is a diagonal matrix of computations parameters.

@@

\begin{thm}[Valid periodic plan]\label{thm:state-vs-energy}  
  Suppose Assumption~\ref{assm:valid}, Lemma~\ref{lem:eqv}, and Proposition~\ref{assm:periodic} hold (a period of the plan is valid, its energy evolution is described by Equation~(\ref{eq:state-perf}), and the plan is periodic). Then the plan is valid and its energy evolution modeled by the periodic energy model in Section~\ref{sec:model}.
\end{thm}
\begin{proof}
  The proof is based on mathematical induction. 
  Base case:

  Induction step:

@@

\subsection{Output constraint set}

@@

\begin{defn}[Output constrain set]
The output constrain set is then the set
\begin{equation*}
  \mathcal{Y}_k:=\{y_k\mid y_k\in[0,\mathrm{SoC}_kQ_cV]\subseteq{\mathbb{R}_{\geq 0}}\},
\end{equation*}
and $\max{\mathcal{Y}}_k$ is the maximum discharge capacity by the internal battery voltage--the maximum instantaneous energy consumption.
\end{defn}
----

@@

\begin{algorithm}
  \begin{algorithmic}[1]
    \STATE $\mathbf{p}_k\gets \mathbf{p}_0, \mathbf{q}_k\gets \mathbf{q}_0$
    \FOR{$i\in[l]^+$}
      \WHILE{$\mathbf{p}_k\neq\mathbf{p}_{\Gamma_i}$}
        \STATE $\mathbf{u}_k\gets\argmax_{\mathbf{u}}\hspace*{-1ex}{\displaystyle{\sum\limits_{i=k-1}^{k+N-2}}\hspace*{-1ex}{l(\mathbf{q}_i,\mathbf{u}_i)+V_f(\mathbf{q}_{k+N-1})}}$\label{alg:mpc}
        \STATE $\mathbf{u}\gets\{\mathbf{u}\cup\mathbf{u}_k\}$
        \STATE $\hat{\mathbf{q}}_{k+1}\gets A\mathbf{q}_k+B\mathbf{u}_k$\label{alg:evolution}
        \STATE $\mathbf{q}_{k+1}\gets$ the estimate of the state from $\hat{\mathbf{q}}_{k+1}$ and sensor data\label{alg:kalman_end}
        \STATE $\mathbf{p}_{k}\gets\mathbf{p}_k\dot{\mathbf{p}}_d(\mathbf{p}_k)/v$ \label{alg:pos}
        \STATE $k\gets k+1$
      \ENDWHILE
    \ENDFOR
  \end{algorithmic}
  \caption{Energy-Aware Dynamic Planning}    
\end{algorithm}

Per each time step $k$ (the final time $t_f$ is unknown), the algorithm updates the state--the position at line~\ref{alg:pos} and the energy coefficients at line~\ref{alg:en}--and the control input. Note that the position can be computed directly from Equation~(\ref{eq:pd}). If the velocity is $v\in\mathbb{R}_{\geq 0}$, and the starting point $\mathbf{p}_0$, $\mathbf{p}_{k+1}=\mathbf{p}_k\dot{\mathbf{p}}_d(\mathbf{p}_k)/v$.

In detail, initial guess for $P_0\in\mathbb{R}^{j\times j}$ is positive definite and derived empirically, for $\mathbf{q}_0$ the initial measurement is distributed to the coefficients (see Theorem~\ref{thm:state-vs-energy}). Line~\ref{alg:max_cont_sequence} selects the maximum possible control from the current control input. Line~\ref{alg:mpc} uses robust output feedback model predictive control (MPC)~\cite{rawlings2017model} to select the optimal control $\mathbf{u}^0$ for a given horizon $N\in\mathbb{Z}_{>0}$ from the cost function
\begin{equation}\begin{split}
  l(\mathbf{q}_k,\mathbf{u}_k)&:=(1/2)(\mathbf{q}_k^TQ\mathbf{q}_k+\mathbf{u}_k^TR\mathbf{u}_k),\\
  V_f(\mathbf{q}_k)&:=(1/2)(\mathbf{q}_k^TP_f\mathbf{q}_k),
\end{split}\end{equation}
where matrices $Q\in\mathbb{R}^{j\times j},R\in\mathbb{R}^{l\times l}$ are positive definite.

Follows a check if the mission can finish without the eventuality of battery discharge (output constraint satisfaction) at line~\ref{alg:output_constraints}, with the control input being eventually updated and the process reiterated at line~\ref{alg:recursion}.

Before the next step, state estimator--the discrete-time Kalman filter~\cite{simon2006optimal} at lines~\ref{alg:kalman_start}--\ref{alg:kalman_end}--predicts the state $\mathbf{q}$ if the modeled instantaneous energy consumption diverges from the sensor's value $y_k^a$ more than a given $\varepsilon\in\mathbb{R}_{\geq  0}$, or sensor measurements are unavailable ($y_k^a=0$).

%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:experimental}

@@

%Some preliminary results are discussed. Further analysis and re-implementation of the functionality of the algorithm is needed.

%\begin{figure}[h]
%  \centering
%  \input{\figpath/trajectory/trajectory3.tikz}
%  \caption{The true trajectory. The trajectory is retrieved from the flight log.}
%  \label{fig:trajectory}
%\end{figure}

%\begin{figure}[h]
%  \centering
%  \input{\figpath/energy/max_qos_tees3.tikz}
%  \caption{Derived energy consumption from the sensors (throttle). Red line is the estimated value. We use KF for state estimation when the difference between the sensor and model value is $>\varepsilon$. Otherwise we evolve the state without noise.}
%  \label{fig:energy-max}
%\end{figure}

@@

Once the UAV reaches the triggering point (line~\ref{code:trigger}) and there are no further lines, the algorithm terminates (the last triggering point is the final point). If there are further lines, the stage switches to $\Gamma_2$
\begin{algorithmic}[1]
  \setcounterref{ALC@line}{code:qos2set}
  \STATE\textbf{\texttt{0.05;1;x+sqrt(4900+c1)+45}}\label{code:tee2}
  \STATE\textbf{\texttt{-sqrt(4900+c1)-45,11}}\label{code:trigger2}
  \STATE\textbf{\texttt{[-3*10\^{}3,0]}}\label{code:teeset2}
  \STATE\textbf{\texttt{[2,10]}}
  \STATE\textbf{\texttt{[32,448]}}\label{code:qosset12}
\end{algorithmic}
line~\ref{code:tee2} corresponds to the trajectory $\varphi_2(\mathbf{p}_k,c_{1,1}):=x+\sqrt{4900+c_{1,1}}+45$, a linear equation that intersects the triggering point (line~\ref{code:trigger}). The gain is $k_e=5\cdot 10^{-2}$, and the rotation is opposite to line~\ref{code:firstcircle}. The algorithm computes $-E$ from the rotation matrix in Equation~\ref{eq:pd}.

>>>>

Let us assume that the set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathcal{C}_i\},
\end{equation}
delimits the area where the $i$-th trajectory $\varphi_i$ is free to evolve using the trajectory parameters $c_{i,1},...,c_{i,\rho}$ (the gray area in Figure~\ref{fig:tee1}). $\varphi_i$ is a function of the two coordinates and is equal to zero when a point $\mathbf{p}_k$ is on the trajectory. Physically, this means the UAV is flying exactly over the nominal trajectory. The trajectory parameters allows to change the trajectory. They are a way to alter the nominal trajectory in the initial plan and thus alter the energy signal.
In fact, the algorithm uses the set from Equation~(\ref{eq:area}) to find the optimal trajectory parameters $c_i^0$, such that $\varphi_i$ has the highest instantaneous energy consumption, while still respecting the energy budget. In Figure~\ref{fig:tee1}, the parameter radius of the circle $c_{1,1}$ is replanned (reduced to $c_{1,1}^-$), as e.g., averse atmospheric conditions do not allow to terminate the plan. The sequence of these parameters (control sequence) forms the path $\mathcal{P}$. The gain factors for the trajectory parameters from Equation~(\ref{eq:state-control}) are obtained
\begin{equation}\label{eq:traj-gains}
  \begin{bmatrix}\nu_{i,1} \\ \vdots \\ \nu_{i,\rho}\end{bmatrix}=(\overline{c}_i-\underline{c}_i)^{-1}\rho^{-1}y_{\underline{c}_i}\begin{bmatrix}c_{i,1} \\ \vdots \\ c_{i,\rho}\end{bmatrix},
\end{equation}
where $y_{\underline{c}_i}\in\mathbb{R}$ is the maximum energy consumption of the highest possible parameters level at $i$-th stage. The value is obtained empirically (e.g., using an energy sensor). Whenever the trajectory parameters are not equally distributed, one can define $y_{\underline{c}_i}$ as a diagonal matrix where the items of the diagonal are the highest possible levels of specific trajectory parameters.  Moreover, let $\nu_{i,1},\dots,\nu_{i,\rho}$ be zero when the parameters $c_i^\rho=\{\emptyset\}$.

We derive the new position $\mathbf{p}_{k+1}$ computing the vector field $\nabla\varphi_i:=\begin{bmatrix}\partial\varphi_i/\partial x & \partial\varphi_i/\partial y\end{bmatrix}^T$, and the direction to follow in the form of velocity vector~\cite{de2017guidance}
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varphi_i-k_e\varphi_i\nabla\varphi_i,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $E$ specifies the rotation (it influence the tracking direction), and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading $\dot{\mathbf{p}}$ due to the atmospheric interferences, such as wind ($w\in\mathbb{R}$ in the top of Figure~\ref{fig:tee1}).

@@

Let us define $g:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ as the instantaneous computational energy consumption value obtained using the tool.

The gain factors that quantifies the contribution of the computations parameters
\begin{equation}\label{eq:energy-comp}
  \begin{bmatrix}\nu_{i,\rho+1} \\ \vdots \\ \nu_{i,\rho+\sigma}\end{bmatrix}=\begin{bmatrix}c_{i,\rho+1}^{-1} & & \\ & \ddots & \\ & & c_{i,\rho+\sigma}^{-1} \end{bmatrix}\begin{bmatrix}g(c_{i,\rho+1}) \\ \vdots \\ g(c_{i,\rho+\sigma})\end{bmatrix},
\end{equation}
where the matrix is a diagonal matrix of computations parameters. Moreover, let $\nu_{i,\rho+1},\dots,\nu_{i,\rho+\sigma}$ be zero when the parameters $c_i^\sigma=\{\emptyset\}$.

@@

\begin{thm}[Valid periodic plan]\label{thm:state-vs-energy}  
  Suppose Assumption~\ref{assm:valid}, Lemma~\ref{lem:eqv}, and Proposition~\ref{assm:periodic} hold (a period of the plan is valid, its energy evolution is described by Equation~(\ref{eq:state-perf}), and the plan is periodic). Then the plan is valid and its energy evolution modeled by the periodic energy model in Section~\ref{sec:model}.
\end{thm}
\begin{proof}
  The proof is based on mathematical induction. 
  Base case: *

  Induction step: *

@@

\subsection{Output and control constraint sets}

@@

\begin{defn}[Output constrain set]
The output constrain set is then the set
\begin{equation*}
  \mathcal{Y}_k:=\{y_k\mid y_k\in[0,\mathrm{SoC}_kQ_cV]\subseteq{\mathbb{R}_{\geq 0}}\},
\end{equation*}
and $\max{\mathcal{Y}}_k$ is the maximum discharge capacity by the internal battery voltage--the maximum instantaneous energy consumption.

The control constraint set is the trajectory contraint set for the trajectory parameters and computation constraint sets for the computation parameters (Definition~\ref{def:mission})
\begin{equation*}
  \mathcal{U}_k:=\begin{cases}
    \mathcal{C}_i & \text{for } c_{i,j} \text{ with } j\leq\rho\\
    \mathcal{S}_{i,j-\rho} & \text{for } c_{i,j} \text{ with } \rho<j\leq\sigma
  \end{cases}.
\end{equation*}
\end{defn}

@@

\begin{algorithm}[t]
  \caption{Energy-Aware Dynamic Planning}\label{alg}
  \begin{algorithmic}[1]
    \STATE $\mathbf{p}_{k-1}\gets \mathbf{p}_0, \mathbf{q}_{k-1}\gets \mathbf{q}_0, \mathbf{u}_{k-1}\gets \{\emptyset\}$\label{alg:init}
    \FOR{$i\in[l]^+$}\label{alg:stages-loop}
      \WHILE{$\mathbf{p}_{k-1}\neq\mathbf{p}_{\Gamma_i}$}
        \STATE $\mathbf{u}_k\gets\argmax_{\mathbf{u}}\hspace*{-1ex}{\displaystyle{\sum\limits_{j=k-1}^{k+N-2}}\hspace*{-1ex}{l(\mathbf{q}_j,\mathbf{u}_j)+V_f(\mathbf{q}_{k+N-1})}}$\label{alg:mpc}
        \STATE $\hat{\mathbf{q}}_{k}\gets A\mathbf{q}_{k-1}+B_i(\mathbf{u}_{k}-\mathbf{u}_{k-1})$\label{alg:evolution}
        \STATE $\mathbf{q}_{k}\gets$ the estimate of the state from $\hat{\mathbf{q}}_{k}$ and sensor data\label{alg:kalman_end}
        \STATE $\mathbf{p}_{k}\gets\mathbf{p}_{k-1}\dot{\mathbf{p}}_d(\mathbf{p}_{k-1})/v$ \label{alg:pos}
        \STATE $k\gets k+1$
      \ENDWHILE
    \ENDFOR
  \end{algorithmic} 
\end{algorithm}

The algorithm first initializes the position, energy coefficients, and control (line~\ref{alg:init}).

The position is updated at line~\ref{alg:pos}, using the expression from Equation~(\ref{eq:pd}) and the velocity $v\in\mathbb{R}_{\geq 0}$. The expression depends on the trajectory $\varphi_i$ from stage $\Gamma_i$. The algorithm iterates all the stages in the plan $\Gamma$ (line~\ref{alg:stages-loop}), and enters the next stage $\Gamma_{i+1}$ when the UAV reaches the triggering point $\mathbf{p}_{\Gamma_i}$.

The energy coefficients are updated at line~\ref{alg:evolution}, using the expression from Equation~(\ref{eq:state-perf}). The a priori state estimate $\hat{\mathbf{q}}_k$ is refined using a state estimator--such as Kalman filter~\cite{simon2006optimal}--and the data from an energy sensor (line~\ref{alg:kalman_end}).
At line~\ref{alg:mpc}, the algorithm uses robust output feedback model predictive control (MPC)~\cite{rawlings2017model} to select the control $\mathbf{u}_k$ for a given horizon $N\in\mathbb{Z}_{>0}$ from the cost function
\begin{equation}\label{eq:cost}\begin{split}
  l(\mathbf{q}_k,\mathbf{u}_k)&:=(1/2)(\mathbf{q}_k^T\mathrm{diag}(C)\mathbf{q}_k)+\\
  &\hspace*{3.5ex}(1/2)(\mathbf{u}_k-\mathbf{u}_{k-1})^TB_i(\mathbf{u}_k-\mathbf{u}_{k-1}),\\
  V_f(\mathbf{q}_k)&:=(1/2)(\mathbf{q}_k^TA\mathbf{q}_k),
\end{split}\end{equation}
where $\mathrm{diag}(C)$ is a diagonal matrix with the items of the matrix $C$ from Equation~(\ref{eq:state-details}) on the diagonal.

We note from Lemma~\ref{lem:eqv} that the expression in Equation~(\ref{eq:cost}) denotes the square of the predicted instantaneous energy consumption. Moreover, higher the horizon, higher the complexity and the accuracy of the control.

We further note that at every step of the summary on line~\ref{alg:mpc} the algorithm has to evolve the state to check if the output satisfies the output constraint set, and if the control satisfies the trajectory and the computation constraint sets. In particular, at each steps it performs the subroutine
\begin{algorithmic}
  {\let\fmtlinenumber\mathalph 
  \WHILE{$\overline{c}_i\notin\mathcal{U}_k,\, y_k\notin\mathcal{Y}_k$}\label{alg2:while}
    \STATE $\overline{c}_i\gets\overline{c}_i-\delta$
  \ENDWHILE
  \STATE $\mathbf{u}_k\gets\overline{c}_i$}
\end{algorithmic} 
where to evaluate the inclusion in the output control set the algorithm evolves the state from its previous estimate similarly as at line~\ref{alg:evolution} (we note that the condition can be written $\overline{c}_i\notin\mathcal{U}_k,\, C\mathbf{q}_k\notin\mathcal{Y}_k$).

We finally note that one can express the tradeoffs between trajectory and computation parameters (e.g., a decrement in the distance between the lines in a survey scenario is related to a decrement in the number of detections per second) enriching the control constraint set with
\begin{equation*}
  F_i\mathbf{u}_k-e_i\geq 0,
\end{equation*}
where $e_i\in\mathbb{R}^n$ and $F_i\in\mathbb{R}^{n\times n}$ expresses the relation between the parameters (if $F_i$ is an identity matrix there is no relation between the parameters).

%%%%%%%%%%%%%%%%%%%%

\section{Results}
\label{sec:experimental}

@@

@@

Once the UAV reaches the triggering point (line~\ref{code:trigger}) and there are no further lines, the algorithm terminates (the last triggering point is the final point). If there are further lines, the stage switches to $\Gamma_2$
\begin{algorithmic}[1]
  \setcounterref{ALC@line}{code:qos2set}
  \STATE\textbf{\texttt{0.05;1;x+sqrt(4900+c1)+45}}\label{code:tee2}
  \STATE\textbf{\texttt{-sqrt(4900+c1)-45,11}}\label{code:trigger2}
  \STATE\textbf{\texttt{[-3*10\^{}3,0]}}\label{code:teeset2}
  \STATE\textbf{\texttt{[2,10]}}
  \STATE\textbf{\texttt{[32,448]}}\label{code:qosset12}
\end{algorithmic}
line~\ref{code:tee2} corresponds to the trajectory $\varphi_2(\mathbf{p}_k,c_{1,1}):=x+\sqrt{4900+c_{1,1}}+45$, a linear equation that intersects the triggering point (line~\ref{code:trigger}). The gain is $k_e=5\cdot 10^{-2}$, and the rotation is opposite to line~\ref{code:firstcircle}. The algorithm computes $-E$ from the rotation matrix in Equation~(\ref{eq:pd}).
